### 2021-05-19 paper/paper_panarchy.Rmd
revised version addressing reviewer comments, plus typo correction on figures

---
title: "Panarchy: ripples of a boundary concept"
author: |
    | \small Juan Rocha^1,2,3^, Linda Luvuno^4^, Jesse Rieb^5^, Erin Crockett^6^,  Katja Malmborg^1^, Michael Schoon^7^, Garry Peterson^1^
    | \small
    | \footnotesize ^1^Stockholm Resilience Centre, Stockholm University, Kräftriket 2B, 10691 Stockholm
    | \footnotesize ^2^Future Earth, The Swedish Royal Academy of Sciences, Lilla Frescativägen 4A, 104 05 Stockholm
    | \footnotesize ^3^South American Institute for Resilience and Sustainability Studies, Maldonado, Uruguay
    | \footnotesize ^4^Center for Complex Systems in Transition, Stellenbosch University, South Africa
    | \footnotesize ^5^Department of Geography, McGill University, 805 rue Sherbrooke ouest, Montréal (Québec) H3A 0B9, Canada
    | \footnotesize ^6^Department of Natural Resource Sciences, McGill University, 21111 Lakeshore Road, 
    | \footnotesize Sainte-11 Anne-de-Bellevue, QC, H9X 3V9, Canada
    | \footnotesize ^7^School of Sustainability, Arizona State University, Tempe, Arizona, USA
abstract: |
  How do social-ecological systems change over time? In 2002 Holling and colleagues proposed the concept of Panarchy, which presented social-ecological systems as an interacting set of adaptive cycles, each of which is produced by the dynamic tensions between novelty and efficiency at multiple scales. Initially introduced as a conceptual framework and set of metaphors, panarchy has gained the attention of scholars across many disciplines and its ideas continue to inspire further conceptual developments. Almost twenty years after this concept was introduced we review how it has been used, tested, extended and revised.  We do this by combining qualitative methods and machine learning. Document analysis was used to code panarchy features that are commonly used in the scientific literature (N = 42), a qualitative analysis that was complemented with topic modeling of 2177 documents. We find that the adaptive cycle is the feature of panarchy that has attracted the most attention. Challenges remain in empirically grounding the metaphor, but recent theoretical and empirical work offer some avenues for future research.
bibliography: panarchy.bib
# csl: nature.csl
output:
  pdf_document:
    dev: pdf
    keep_tex: yes
    toc: no
    latex_engine: xelatex
  rtf_document:
    toc: no
citecolor: blue
documentclass: article
font: Arial
fontsize: 10pt
linkcolor: blue
urlcolor: blue
header-includes:
- \usepackage{dcolumn, rotating, longtable, lineno, float, fontspec, array, tabularx}
- \setmainfont{Arial}
- \renewcommand{\abstractname}{}
- \graphicspath{{figures/}}
- \usepackage[margin=2.5cm]{geometry}

## For Arxiv submission, use pdflatex as engine, comment out font, increase font size to 11, delete the package fontspec and the graphics path. Compile in latex so the bibliography is part of the tex file produced. 
---

# Introduction

Almost two decades ago, the edited book "Panarchy: Understanding Transformations
in Human and Natural Systems" [@Gunderson:2002vk], presented a synthetic
perspective of how a group of social-ecological researchers associated with the
Resilience Alliance understood change in social-ecological systems. The concept
of Panarchy was a key focus on this influential book. In the late 1990s and
early 2000s, the Resilience Alliance was a productive, innovative, and highly
collaborative group of interdisciplinary scientists that focused on bridging
social and natural sciences as well as theory and practice. They were addressing
social-ecological problems by combining insights from the social and natural
sciences, as well as the arts and humanities [@Parker:2012eo].

The Panarchy concept builds on Holling's adaptive cycle [@biosphere:hJ4c_Ua8] by
extending the idea across spatial and temporal scales. It remains a boundary
object that has inspired research topics, enabled collaborations and nurtured
new scientific frameworks. The ideas put forward have been applied in field
studies, archaeology, mathematical models, participatory work, and scenario
development. Panarchy has inspired resilience assessments and guided decision
making. In this article we pay tribute to the book by studying how the concepts
and metaphors proposed have been further developed in the academic literature.
We also document criticism of the concepts and identify key research frontiers.

Panarchy proposes that it is useful to conceptualize systems in terms of
interacting adaptive cycles. The adaptive cycle was an idea first proposed by
Holling based upon his experience working and studying managed ecosystems
[@biosphere:hJ4c_Ua8]. It was meant to be a conceptual tool that focused
attention on processes of destruction and reorganization, which had been
neglected in comparison to those of growth and conservation. Considering these
processes provides a more complex way to understand system dynamics.

An adaptive cycle alternates between long periods of system aggregation,
connection, and accumulation, and shorter periods of disruption and
reorganization. The adaptive cycle exhibits four phases arranged into two major
loops (Fig. \ref{fig:fig1}). The phases are exploitation ($r$), conservation
($K$), release ($\Omega$), and reorganization ($\alpha$), the latter
characterized by events of creative destruction (Fig. \ref{fig:fig1}). The first
loop, often referred to as the front-loop, from $r$ to $K$, is the slow,
incremental phase of growth and accumulation. The second loop, referred to as
the back-loop, from $\Omega$ to $\alpha$, is the rapid phase of reorganization
leading to renewal. A system going into the back-loop can either remain in a
similar form, or may transition to a new type of system with new boundaries and
key components [@Holling:2002wu]. Resilience researchers proposed that tensions
between demands for organization and efficiency versus demands for novelty and
diversity drive adaptive cycle type dynamics in many different types of complex
systems [@Holling:2002wu].

Panarchy is not a theory of what it is, but a metaphor of what might be
[@Gunderson:2002vk]. It is not a predictive tool, but aims to understand
adaptive change. The adaptive cycle is hypothesized to exist within a
three-dimensional space defined by three properties: potential, connectivity,
and resilience [@Gunderson:2002vk]. Potential refers to the capital available to
the system, for example nutrients and carbon captured by a forest, or human
capital - skills and knowledge accumulated to run the economy. Connectivity is a
proxy of the structure of the system; it is the network of interactions and
strength between its elements. Resilience is the capacity of any system to
absorb disturbance and reorganize while undergoing change so as to retain
essentially the same identity, as determined by its function, structure, and
feedbacks [@Anonymous:2016fv].

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 2.5in]{fig1_panarchy}
\caption{\textbf{Panarchy} is an heuristic of nested adaptive cycles that serves to represent a variety of systems and environmental problems. Adaptive cycles (A) at different scales of the hierarchy (B) can be connected through remember and revolt cross-scale interactions.}
\label{fig:fig1}
\end{figure*}
```
Panarchy posits that systems are organized in nested hierarchies across space
and time, where each level of the hierarchy is a subsystem that can be in a
different phase of the adaptive cycle. These phases are inspired by mathematical
models used in economics and ecology, but broadly describe patterns of growth,
collapse, and recovery that are common to populations, ecological communities,
markets, or political organizations. While each subsystem in the hierarchy can
be at a different phase of the adaptive cycle, these subsystems can influence
each other through cross-scale interactions called *"revolt"* and *"remember"*
[@Gunderson:2002vk].

The three-dimensional space has corners with attractors that can cause the
adaptive cycle to stagnate: poverty and rigidity traps [@Gunderson:2002vk].
Poverty traps are described in panarchy as maladaptive states where potential,
connectivity and resilience are low. Poverty traps are a series of feedback
mechanisms that reinforce impoverished states [@Bowles:2006hx; @Maru:2012dm;
@Allison:2004te], limiting the system's capacity to innovate and increase
potential. The opposite corner, where potential, connectivity and resilience are
high, is another maladaptive space called a rigidity trap. In that corner there
is little space for experimentation and innovation. Examples include systems
where ecological resilience has been expensively replaced by artificial
processes to maintain the system such as levees, flood barriers, or chemical
control of pests [@Holling:1996wx].

Panarchy offers a rich conceptual framework for understanding environmental
problems. While inspired by a few mathematical constructs (e.g. cycles, traps,
scaling laws), it is general enough to invite scholars from multiple
disciplinary backgrounds, ontologies and epistemologies, to collaborate around
research questions and applied problems. As such, it is useful as a boundary
object and can be used empirically or metaphorically. The book itself presents a
series of case studies where geographers, economists, political scientists, and
ecologists have demonstrated the utility of the framework to their area of
research. However, the extent to which these concepts have gained empirical
support beyond the metaphor, or how they have evolved over time, remain open
questions. For example, the adaptive cycle is a useful metaphor to look back at
history and organize events and periods around phases, but it remains
challenging to identify the phase of a system in real time, or to use panarchy
to project future trajectories.

Here we review the academic literature of the last two decades to trace how
these concepts have evolved, in what type of problems they have been applied and
found useful, and finally what remains as the key frontiers of research.

# Methods

To answer these questions we combined an automated literature review based on
topic modeling [@Blei:2012dk; @Griffiths:2004ey] with human-coded document
analysis [@Bryman:2008va].

*Data*: We used the Web of Science, Scopus, and GoogleScholar to survey academic
literature that has used or referenced works that trace back to Panarchy
[@Gunderson:2002vk]. We extracted complete records from the Scopus database that
matched the search for *"panarchy OR adaptive cycle"* (N = 595), or the search
*"(panarchy OR adaptive cycle) AND resilience"* (N = 278). The data was combined
with all papers (N = 1923) that cited the inaugural paper that introduced the
book to the academic community [@Holling:2001cu]. Records with missing abstracts
were dropped (N = 191), and records with missing year were set to 2020 given
that they are accepted manuscripts with digital object identifier (doi)
scheduled to be published later in 2020 or 2021.

To prepare the data for topic modeling, we constructed a document term matrix
with documents in rows and words in columns. Here our unit of analysis for the
document are the abstracts retrieved, and the matrix contains the count of words
per document. We removed stop words (e.g. "the", "a") and digits from the
matrix, as well as a list of words that were over represented in our data and
are common in the scientific literature but are unrelated to the papers' topics
(e.g. "paper", "study", "aim").

*Topic models*: are an unsupervised statistical technique to reduce the
dimensionality of a corpus of data (typically but not necessarily text) into
topics [@Blei:2012dk]. Here a topic is a latent variable that ranks words with
high probability of appearing together within the same document. Documents in
turn can be described by the probability distribution of a particular set of
topics. Since they are (posterior) probabilities, the sum of the probability of
all words for any given topic should be one, and the sum of the probability of
all topics for any given document should also be one. An iterative process or
algorithm is what allows the model to learn the ranking of words that best
explain topics, and the ranking of topics that best explain documents.

The underlying statistical technique for this machine learning approach is
called Latent Dirichlet Allocation (LDA) [@Blei:2003tn]. It allocates
probabilities to latent variables (topics) based on the distribution of words in
text data, assuming a multivariate continuous (Dirichlet) distribution. We
compared three LDA algorithms: correlated topic models (CTM), variational
expectation maximization (VEM), and Gibbs sampling (Gibbs), by assessing their
performance against their log-likelihood estimation, entropy and perplexity
[@Grun:2011tb]. Entropy is a measure of order or disorder of a system; in the
context of topic models it measures how evenly the topic distribution is spread.
Perplexity measures the uncertainty of predicting a single word, so if the model
performance were the same as random, perplexity would approximate the vocabulary
size (N = 16 728 words). These performance metrics enable us to choose which
algorithm best fits our data, what is the optimal number of topics to fit, and
how to avoid overfitting.

*Document analysis*: We complemented our unsupervised approach with document
analysis [@Bryman:2008va] by coding an additional set of categorical variables
for a sample of papers (N = 42). Based on the reading of Panarchy, we first
developed a series of categorical variables that captured the main components of
the scientific framework. The variables were presented and received feedback at
the science meeting of the Resilience Alliance in 2019. We used the variables to
annotate qualitative aspects such as use of the adaptive cycle, identification
of its phases, whether the paper is conceptual, modeling or an empirical study.
When empirical, we recorded the temporal and spatial scales of the case study.
We also identified what aspects of panarchy were most used in the papers. For
example whether there is emphasis on cross-scale interactions, or poverty and
rigidity traps. We used text annotations to capture potential criticisms as well
as the methods used.

# Results

Panarchy, the book published in 2002, has been cited over 7200 times according
to Google Scholar. The scientific paper that introduced the book to the
scientific community [@Holling:2001cu] has received, at the time of writing,
1715 citations in the Web of Science, and 1923 in Scopus. Roughly half of the
citations have come from environmental (28%) and social sciences (22%). Computer
science (2.1%) and arts and humanities (2.6%) have been the least represented.

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 5in, height = 4in]{fig2_topicnumber}
\caption{\textbf{Algorithm and number of topic selection} Gibbs sampling maximizes entropy and log-likelihood estimation, making it a suitable algorithm for our data (A). Increasing the number of topics (from 5-100) shows that $\alpha$ decreases, suggesting that despite the larger number of topics, a few of them suffice to describe most papers (B). Log-likelihood is maximized for 50 topics followed closely with 25. While 50 topics is marginally better, for visualization purposes we restrict our analysis to 25. Note that perplexity cannot be calculated for Gibbs sampling, hence the missing value in A and absence in B.}
\label{fig:algorithm}
\end{figure*}
```
Comparing algorithm performance revealed that Gibbs sampling was the best fit
for the data (Fig \ref{fig:algorithm}). As a rule of thumb, an ideal method
should maximize entropy and the log-likelihood estimation while minimizing
perplexity [@Grun:2011tb]. Gibbs sampling maximised both entropy and likelihood
with our data when compared to other alternatives. The second best was the
variational-expectation maximisation algorithm (VEM) when $\alpha$ was not set
constant. $\alpha$ is a hyper parameter that weights the evenness of topic
distribution. A lower $\alpha$ than default values indicates that the documents
can be described by rather fewer topics, or that its distribution is very
uneven. In fact, we observed that increasing the number of topics from 5 to 100
topics did increase entropy at expense of reducing $\alpha$, meaning that
despite the larger number of topics, the main content of a document is still
captured by a few of them. The log-likelihood maximization stopped at around
25-50 topics. Thus, we restricted the rest of our analysis to 25 topics.

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 5in, height = 5in]{fig3_topics25}
\caption{\textbf{Panarchy topics over time} The number of papers per year (A) with a maximum of 324 in 2019. The proportion of papers per year (B) and the relative proportion of topic content per year (C) do not show strong trends for the time window with most papers (gray area in A). Each topic is summarized in (D) by the top 10-words that best describe them according to the posterior probability of our model fit. See FigS1 for a panel of time series per topics}
\label{fig:topics25}
\end{figure*}
```
A topic is a set of words that are ranked according to a probability that they
represent an underlying content of the document (Fig \ref{fig:topics25}). For
example, the words "resilience", "vulnerability", "risk" and "disaster" have a
high probability to capture the content of topic 25. Papers early in our time
series (2001-3) have high content largely dominated by topics 25 on resilience,
24 on the adaptive cycle, and 18 on adaptive governance and management. Towards
the end of the time series, topic 3 on urban systems and 9 on local communities
and knowledge have spikes up to 9% of the content of each year (2019-2020). For
comparison, if all topics were equally represented in the content, they would
have 4% in the corpus (grey line in Fig \ref{fig:topics25}B).

Despite fluctuations, most topics showed a relatively constant level of interest
over time (Fig \ref{fig:topics25}, Fig \ref{fig:figS1}). We did not observe
strong trends, but some topics have gained a small amount of attention. For
example, topic 1 on sustainability, or topic 10 on ecosystem services appeared
consistently across time. In contrast, research on business innovation (topic
20), urban infrastructure (topic 3), or archaeology (topic 5) have gained
attention in recent years. Topic 8 was an outlier, with a selection of papers
ranking high on content related to network infrastructure and performance,
possibly from engineering disciplines. It was the only topic with a set of
papers that was clearly distinct from the rest of the collection, and showed a
decreasing trend over time.

The human-coded document analysis revealed that the most common feature of
panarchy in the literature was the adaptive cycle (81%, N = 42) followed by
cross-scale interactions (Fig \ref{fig:qual}). Poverty and rigidity traps were
less studied features in our sample (14% and 19% respectively), even when
accounting for slightly different terminology such as "lock-ins". The bulk of
the papers analyzed were conceptual papers, many without a method or discernible
research question. Roughly half of the papers were conceptual or only used
panarchy as a metaphor. Over 40% of papers in our sample went a step further and
used panarchy as an empirical construct, for example by attempting to identify
the phases of the adaptive cycle (61%). Out of the 22 empirical cases, 6 were at
the time scale of centuries, 7 on decades, 6 on years, and 1 in weeks.
Spatially, 4 were at a city scale, 17 regional, and 1 national. Empirical papers
were dominated by qualitative methods (77%, N = 22) and were generally
retrospective historical reviews.

Several of the papers reviewed attempted to identify the adaptive cycle either
through qualitative or quantitative methods. Identification of adaptive cycles
has found applications in a wide range of disciplines and research problems,
from delimitation of periods in archaeology and anthropology [@Redman:2003bd],
to financial crises in Europe [@Castell:2020dg], or traffic jams in China
[@Zeng:2020hu]. Recent empirical tests of the adaptive cycle innovate on the
types of data and methods that one can use. Information transfer methods based
on entropy have been suggested to approximate relevant components of a system
and the empirical proxies of the axis where the adaptive cycle unfolds:
potential, connectedness, and resilience [@Castell:2020dg]. Percolation methods
combined with big data have been shown to be useful to test hypotheses of regime
shifts in urban systems (originally proposed in Panarchy [@Gunderson:2002vk]),
and derive the temporal and spatial scales at which the adaptive cycle emerges
[@Zeng:2020hu].

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 5in, height = 2.5in]{fig4_qual_results}
\caption{\textbf{Qualitative results} Document analysis was used to disentangle the different panarchy features addressed by a sub-sample of papers (N = 42). Over half of the papers are conceptual work, and most empirical papers fall into qualitative methods.}
\label{fig:qual}
\end{figure*}
```
# Discussion

The high level of conceptualization but lack of empirically grounded theory
testing came across as one of the major limitations. Here we do not refer only
to positivism as a plausible way of theory development. Other equally valid
approaches exist. From the philosophical point of view of what makes a
theory a scientific theory, the plurality of approaches should have mechanisms
in place to avoid confirmation bias, allow the emergence of alternative
explanations, and identify the circumstances under which a theory may fail
[@Hansson:2017tr]. Many of the papers we read confirm but not challenged the
expectations that Panarchy put forward. A few of the most recent work actually
falsify some of the ideas proposed --e.g. the back loop is faster that the front
one. We believe more work is needed to better understand the circumstances where
Panarchy works, and when it falls short. In other words, refine the scientific
framework towards a scientific theory.

Given the strong dominance of conceptual papers, it is not surprising that one
of the key frontiers of research is to empirically ground such
conceptualizations in a way that one can test theoretical expectations beyond
metaphors. Metaphors have played an important role as boundary objects enabling
interdisciplinary dialogues. We observe it on the wide applications that
panarchy has inspired, from anthropology to engineering. In the empirical realm,
the adaptive cycle is the panarchy dimension that researchers engage the most
with, but it is dominated by qualitative and retrospective studies. That means,
we still lack the theory and methods to gather observations and be able to
decide in which phase of the adaptive cycle a system currently is, or what is
the probability of it transitioning to a new phase?

The criticism of over-conceptualization is not unique to the literature engaging
Panarchy ideas. Here by over-conceptualization we refer to the problem of
scientific jargon, where different communities use slightly different wordings
to refer to the same phenomena; reducing the chances that lessons from a
disciplines percolates to another. An example from Panarchy is the endorsement
of ideas such as creative destruction (from economics) or the theory of
intermediate disturbances (from ecology). The latter has been debunked in the
ecological literature but might still been used within Panarchy thinking
[@Fox:2013br]. A recent review of sustainability science mapped the different
schools of thought that the discipline has developed over the last decades
[@Clark:2020jd]. A similar conclusion was reached, where too many conceptual
frameworks have been developed, but fewer empirical attempts try to test the
frameworks against data to falsify hypotheses. The review emphasizes the problem
of measurement and observation [@Clark:2020jd]. In the context of panarchy,
recent work has developed an information theory based approach that enables the
identification of adaptive cycles [@Castell:2020dg]. The authors identify phases
of the adaptive cycle in the European financial crises and in grassland
ecosystems, but fail to find support for the difference in speed between the
forward and backward loops originally proposed in Panarchy.

The panarchy dimensions that received less attention in our qualitative analysis
are creative destruction, rigidity and poverty traps. This may be at least
partly due to the scope of our data: papers that have cited Panarchy or
Holling's paper [@Gunderson:2002vk; @Holling:2001cu]. Concepts such as creative
destruction and poverty traps precede Panarchy and therefore have been theorized
and empirically grounded outside the panarchy stream of thinking. For example,
the theory of poverty traps dates back to the 1950's in economics, and has
received both theoretical development [@Bowles:2006hx] as well as empirical
grounding [@Banerjee:2015ja; @Banerjee:2012tq] that has enabled researchers and
governments to distinguish what kind of interventions are likely to reduce
poverty. Here we encourage similar empirical efforts in formalizing theory and
empirical support for the adaptive cycles, their cross-scale interactions
(remember and revolt), or how the hierarchical nested nature of complex systems
enhance or erode resilience in social-ecological systems.

This paper aimed to study the use of panarchy related concepts and their
evolution since the publication of the book. Systematic literature reviews often
suffer from the limitation of restricting sample size to a subset of what is
readable on the time frame of a project -- between dozens and maybe hundreds of
papers. Qualitative analysis offers rich insights into the papers but is limited
by sample size. In our analysis, the qualitative insights could be biased
towards a non-random selection of papers that aligned our research interests, or
highly cited, review type of work. Topic modeling enabled us to complement the
analysis to all papers reported in major scientific databases. It has the
advantage of reproducibility and reduces sample bias, but offers little insights
about the dimensions of panarchy used, methods, or criticisms. Here we show that
both methods combined are a powerful combination.

Our choice of methods also has limitations. We could not compare both methods
since they do not have a common categorical axis on which they can be measured
back to back. The topics gave us a general but shallow overview, while the
document analysis provide deep insights on particular work but limited
generalizations. We, for example, ignored rich social science literature (e.g.
Anthropology) that is published in books and book chapters, hence not archived
in scientific search engines. A more inductive approach to the qualitative
coding could have gave us different insights on the literature outside the
pre-defined categories we initially shared with the Resilience Alliance.

Future studies could benefit from including gray literature such as theses,
books, non-governmental organizations reports, government agencies reports, or
non-english literature. Previous studies have shown that the use of full text
instead of abstracts can also offer additional insights on the automated
analyses, for example in attributing impacts of ecosystem services from regime
shifts in social-ecological systems [@Rocha:2015tk]. Replication attempts do
however face the challenge of accessing full text when many of the papers are
behind pay-walls.

# Conclusion

We reviewed the academic literature of the last two decades to trace how
Panarchy related concepts have evolved, in what type of problems they have been
applied and found useful, and what remains as the key frontiers of research.
Despite a growing body of literature, no topic seems to dominate the academic
production of scholars using these concepts. The feature most used is the
adaptive cycle, and the problems where it is found most useful is in studies
with a historical perspective, either in short time horizons such as natural
resource management, urban development, or conservation; all the way to
archeology and anthropology studies on the time horizon of centuries to
millennia. Hierarchies in scale or maladaptive traps have received comparatively
less attention. The frontiers of research point out to the problem of
measurement and prediction. For example, how to observe and approximate
resilience, connectedness and potential in the present; or how likely a current
social-ecological system is to undergo a transition in the phases of the
adaptive cycle, or a "memory" or "revolt" type of cross-scale interaction. While
our survey of the literature offers a few options to empirically ground panarchy
concepts, we believe the problem of measurement and prediction are fertile
ground for future research efforts.

# Acknowledgements

JCR acknowledges financial support from Formas research grant (942-2015-731),
the Stockholm Resilience Centre, and Stockholm University.

# Data availability

All data and code to replicate the analysis are available at:
<https://github.com/juanrocha/panarchy>. The code book and database are
available in thea public online repository
([https://doi.org/10.6084/m9.figshare.13490919](doi.org/10.6084/m9.figshare.13490919)).

# References

::: {#refs}
:::

\pagebreak

# Supplementary Material {#sec:SM}

```{=tex}
\renewcommand\thefigure{S\arabic{figure}}
\renewcommand\thetable{S\arabic{table}}
\setcounter{table}{0}
\setcounter{figure}{0}
```
```{=tex}
\begin{figure*}[h]
\centering
\includegraphics[width = 6in, height = 5in]{SM_topics25_timeline}
\caption{\textbf{Proportion of papers per year per topic.} Each line represent the average proportion of content for a given year over all papers published in the same year.}
\label{fig:figS1}
\end{figure*}
```

```{=tex}
\begin{figure*}[h]
\centering
\includegraphics[width = 6in, height = 5in]{sm2}
\caption{\textbf{Proportion of topics per paper.} Each dot is a paper with its respective proportion of topic per year. A loess fit (blue lines) show the lack of trends over time for most topics. The color of topics correspond to Fig \ref{fig:topics25}}
\label{fig:figS2}
\end{figure*}
```