### 2022-03-07 paper/paper_panarchy.Rmd
implementing suggestions from reviewers and co-authors

---
title: "Panarchy: ripples of a boundary concept"
author: |
    | \small Juan C. Rocha^1,2,3^, Linda Luvuno^4^, Jesse Rieb^5^, Erin Crockett^6^,  Katja Malmborg^1^, Michael Schoon^7^, Garry D. Peterson^1^
    | \small
    | \footnotesize ^1^Stockholm Resilience Centre, Stockholm University, Kräftriket 2B, 10691 Stockholm
    | \footnotesize ^2^Future Earth, The Swedish Royal Academy of Sciences, Lilla Frescativägen 4A, 104 05 Stockholm
    | \footnotesize ^3^South American Institute for Resilience and Sustainability Studies, Maldonado, Uruguay
    | \footnotesize ^4^Center for Complex Systems in Transition, Stellenbosch University, South Africa
    | \footnotesize ^5^Department of Geography, McGill University, 805 rue Sherbrooke ouest, Montréal (Québec) H3A 0B9, Canada
    | \footnotesize ^6^Department of Natural Resource Sciences, McGill University, 21111 Lakeshore Road, 
    | \footnotesize Sainte-11 Anne-de-Bellevue, QC, H9X 3V9, Canada
    | \footnotesize ^7^School of Sustainability, Arizona State University, Tempe, Arizona, USA
abstract: |
  How do social-ecological systems change over time? In 2002 Holling and colleagues proposed the concept of panarchy, which presented social-ecological systems as an interacting set of adaptive cycles, each of which is produced by the dynamic tensions between novelty and efficiency at multiple scales. Initially introduced as a conceptual framework and set of metaphors, panarchy has gained the attention of scholars across many disciplines, and its ideas continue to inspire further conceptual developments. Almost twenty years after this concept was introduced we review how it has been used, tested, extended and revised.  We do this by combining qualitative methods and machine learning. Document analysis was used to code panarchy features that are commonly used in the scientific literature (N = 42), a qualitative analysis that was complemented with topic modeling of 2177 documents. We find that the adaptive cycle is the feature of panarchy that has attracted the most attention. Challenges remain in empirically grounding the metaphor, but recent theoretical and empirical work offer some avenues for future research.
bibliography: [panarchy.bib, packages.bib]
csl: "/Users/juanrocha/Documents/styles/ecology.csl"
output:
  rtf_document:
    toc: no
  pdf_document:
    dev: pdf
    keep_tex: yes
    toc: no
    latex_engine: xelatex
citecolor: blue
documentclass: article
font: Arial
fontsize: 10pt
linkcolor: blue
urlcolor: blue
header-includes:
- \usepackage{dcolumn, rotating, longtable, lineno, float, fontspec, array, tabularx}
- \setmainfont{Arial}
- \renewcommand{\abstractname}{}
- \graphicspath{{figures/}}
- \usepackage[margin=2.5cm]{geometry}
---

```{r message = FALSE, warnings=FALSE, include = FALSE}
library(here)
library(tidytext)
library(tidyverse)
library(tm)
library (topicmodels)
library(lda)

load(here("data/dtm.Rdata"))
load(here("data/qual_coding.Rdata"))
```

# Introduction

Almost two decades ago, the edited book "Panarchy: Understanding Transformations
in Human and Natural Systems" [@Gunderson:2002vk], presented a synthetic
perspective of how a group of social-ecological researchers associated with the
Resilience Alliance understood change in social-ecological systems. The concept
of panarchy was a key focus of this influential book. In the late 1990s and
early 2000s, the Resilience Alliance was a productive, innovative, and highly
collaborative group of interdisciplinary scientists that focused on addressing
social-ecological problems. They did this by combining insights from the social
and natural sciences, arts, and humanities, as well as by bridging theory and
practice [@Parker:2012eo].Panarchy remains a boundary
object that has inspired research topics, enabled collaborations and nurtured
new scientific frameworks [@Parker:2012eo]. The ideas put forward have been applied in field
studies, archaeology, mathematical models, participatory work, and scenario
development [@gunderson2022applied]. Panarchy has inspired resilience assessments and guided decision
making. In this article we pay tribute to the book by studying how the concepts
and metaphors proposed have been further developed in the academic literature.
We also document criticism of the concepts and identify key research frontiers.

The panarchy concept builds on Holling's adaptive cycle [@biosphere:hJ4c_Ua8]
by extending the idea across spatial and temporal scales. Panarchy proposes that it is useful to conceptualize systems in terms of
interacting adaptive cycles. The adaptive cycle was an idea first proposed by
Holling based upon his experience working and studying managed ecosystems
[@biosphere:hJ4c_Ua8]. It was meant to be a conceptual tool that focused
attention on processes of destruction and reorganization, which had been
neglected in comparison to those of growth and conservation.

An adaptive cycle alternates between long periods of system aggregation,
connection, and accumulation, and shorter periods of disruption and
reorganization. The adaptive cycle exhibits four phases arranged into two major
loops (Fig. \ref{fig:fig1}). The phases are exploitation ($r$), conservation
($K$), release ($\Omega$), and reorganization ($\alpha$), the latter
characterized by events of creative destruction (Fig. \ref{fig:fig1}). The first
loop, often referred to as the front-loop, from $r$ to $K$, is the slow,
incremental phase of growth and accumulation. The second loop, referred to as
the back-loop, from $\Omega$ to $\alpha$, is the rapid phase of reorganization
leading to renewal. A system going into the back-loop can either remain in a
similar form, or may transition to a new type of system with new boundaries and
key components [@Holling:2002wu]. Resilience researchers proposed that
tensions between demands for organization and efficiency, versus demands for
novelty and diversity, drive adaptive cycle dynamics in many different types of
complex systems [@Holling:2002wu].

"_Panarchy is not a theory of what it is, but a metaphor of what might be_"
[@Gunderson:2002vk 32]. It is not a predictive tool, but aims to understand
adaptive change. The adaptive cycle is proposed to exist within a
three-dimensional space defined by three properties: potential, connectedness,
and resilience [@Gunderson:2002vk]. Potential refers to the capital available
to the system, for example nutrients and carbon captured by a forest, or human
capital - skills and knowledge accumulated to run the economy. Connectedness is
a proxy of the structure of the system; it is the network of interactions and
strength between its elements. Resilience is the capacity of any system to
absorb disturbance and reorganize while undergoing change so as to retain
essentially the same identity, as determined by its function, structure, and
feedbacks [@Anonymous:2016fv].

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 2in]{fig1_panarchy}
\caption{\textbf{Panarchy} is an heuristic of nested adaptive cycles that serves to represent a variety of systems and environmental problems. Adaptive cycles (A) at different scales of the hierarchy (B) can be connected through remember and revolt cross-scale interactions. (C) relates the adaptive cycle to the potential, connectedness and resilience axis, reproduced from Panarchy edited by Lance Gunderson and C.S. Holling, copyright \textcopyright 2002 Island Press. Reproduced with permission of Island Press, Washington, DC.}
\label{fig:fig1}
\end{figure*}
```
Panarchy posits that systems are organized in nested hierarchies across space
and time, where each level of the hierarchy is a subsystem that can be in a
different phase of the adaptive cycle. These phases are inspired by mathematical
models used in economics and ecology, but broadly describe patterns of growth,
collapse, and recovery that are common to populations, ecological communities,
markets, or political organizations. While each subsystem in the hierarchy can be at a different phase of the adaptive cycle, these subsystems can particularly influence each other through cross-scale interactions called *“revolt”*, where release in a smaller, faster sub-system can trigger release in a larger, slower sub-system,  and *“remember”*, in which the structure of a larger-slower sub-system shapes the dynamics of reorganization of a smaller, faster sub-system
[@Gunderson:2002vk].

The three-dimensional space defined by potential, connectedness and resilience, has corners with attractors that can derail the
adaptive cycle: poverty and rigidity traps [@Gunderson:2002vk].
Poverty traps are described in panarchy as maladaptive states where potential,
connectedness and resilience are low. Poverty traps are a series of feedback
mechanisms that reinforce impoverished states [@Bowles:2006hx; @Maru:2012dm;
@Allison:2004te], limiting the system's capacity to innovate and increase
potential. The opposite corner, where potential, connectivity and resilience are
high, is another maladaptive space called a rigidity trap. In that corner there
is little space for experimentation and innovation. Examples include systems
where ecological resilience has been extensively replaced by artificial
processes to maintain the system such as levees, flood barriers, or chemical
control of pests [@Holling:1996wx].

Panarchy offers a rich conceptual framework for interpreting environmental
problems. While inspired by several mathematical constructs (e.g. cycles, traps,
scaling laws), it is general enough to invite scholars from multiple
disciplinary backgrounds, ontologies and epistemologies, to collaborate around
research questions and applied problems. As such, it is useful as a boundary
object and can be used empirically or metaphorically. The book presents a
series of case studies where geographers, economists, political scientists, and
ecologists have demonstrated the utility of the framework to their area of
research.

Scientific theories are bodies of knowledge that answer the
question why? The answer implies causal reasoning. Regardless of discipline, a
theory should therefore provide a series of testable expectations that help
distinguishing the situations when it is a good explanation of reality, versus
situations when it falls short. The practice of science implies avoiding
confirmation bias and testing when explanations are valid representations of
reality. In this paper we explore how the conceptual framework of panarchy has been applied, developed and tested over the past twenty years. For panarchy to become a useful theory in sustainability science, it 
needs to be able to explain phenomena in
social-ecological systems across different contexts, in a way tah tis testable, and withstands the rigour of those tests. Here we review the academic literature of the last two decades to trace how the
ideas proposed in panarchy have evolved. We conclude by suggesting research frontiers for future work.

# Methods

To answer these questions we combined an automated literature review based on
topic modeling [@Blei:2012dk; @Griffiths:2004ey] with human-coded document
analysis [@Bryman:2008va]. We first describe the data curation and how it was
used in the topic models. Then we introduce the qualitative approach.

*Data*: We used the Web of Science, Scopus, and GoogleScholar to survey academic
literature that has used or referenced works that trace back to panarchy
[@Gunderson:2002vk]. We extracted complete records from the Scopus database
that matched the search for *"panarchy OR adaptive cycle"* (N = 595), or the
search *"(panarchy OR adaptive cycle) AND resilience"* (N = 278). The data was
combined with all papers (N = 1923) that cited the inaugural paper that
introduced the book to the academic community [@Holling:2001cu]. Records with
missing abstracts were dropped (N = 191), and records with missing year were set
to 2020 given that they are accepted manuscripts with digital object identifier
(doi) scheduled to be published later in 2020 or 2021. Our final sample is
`r dim(dtm)[1]` documents.

To prepare the data for topic modeling, we constructed a document term matrix
with documents in rows (N=`r dim(dtm)[1]`) and words in columns
(N=`r dim(dtm)[2]`). Here our unit of analysis for the document are the
abstracts retrieved, and the matrix contains the count of the number of
occurrences of each word per abstract. We removed stop words (e.g. "the", "a")
and digits from the matrix, as well as a list of words that were over
represented in our data and are common in the scientific literature but are
unrelated to the papers' topics (e.g. "paper", "study", "aim"). The lexicon of
stop words has 1149 entries, curated by various research groups, and accessible
in the `R` language under the packages `tm` and `tidytext` [@R-topicmodels;
@R-tm; @R-NLP; @tm2008; @R-tidytext], for example by `tidytext::stop_words`. The
data was also lemmatized using the hash lemmas dataset available at
[\<http://www.lexiconista.com>](http://www.lexiconista.com/). Lemmatization is
the process of simplifying words to their root lemmas (e.g. *teaching* and
*taught* are lemmatized to *teach*).

*Topic models*: are an unsupervised statistical technique to reduce the
dimensionality of a corpus of data (typically but not necessarily text) into
topics [@Blei:2012dk]. Here a topic is a latent variable that ranks words
with high probability of appearing together within the same document. Documents
in turn can be described by the probability distribution of a particular set of
topics. Since they are (posterior) probabilities, the sum of the probability of
all words for any given topic should be one, and the sum of the probability of
all topics for any given document should also be one. An iterative process or
algorithm is what allows the model to learn the ranking of words that best
explain topics, and the ranking of topics that best explain documents.

The underlying statistical technique for this machine learning approach is
called Latent Dirichlet Allocation (LDA) [@Blei:2003tn]. It allocates
probabilities to latent variables (topics) based on the distribution of words in
text data, assuming a multivariate continuous (Dirichlet) distribution. We
compared three LDA algorithms: correlated topic models (CTM), variational
expectation maximization (VEM), and Gibbs sampling (Gibbs), by assessing their
performance against their log-likelihood estimation, entropy and perplexity
[@Grun:2011tb]. Entropy is a measure of order or disorder of a system; in the
context of topic models it measures how evenly the topic distribution is spread.
Perplexity measures the uncertainty of predicting a single word, so if the model
performance were the same as random, perplexity would approximate the vocabulary
size (N = `r dim(dtm)[2]` words). These performance metrics enable us to choose
which algorithm best fits our data, what is the optimal number of topics to fit,
and how to avoid overfitting.

*Document analysis*: We complemented our unsupervised approach with document
analysis [@Bryman:2008va] by coding an additional set of categorical
variables for a sample of papers (N = `r nrow(qual)`). The papers read are part
of the corpus used for topic modelling. We read papers avoiding biasing our
selection to high cited papers, specific journals, older papers or particular
disciplines or methods.

Based on the reading of panarchy, we first developed a series of categorical
variables that captured the main components of the scientific framework. The
variables were presented and received feedback at the science meeting of the
Resilience Alliance in 2019. We used the variables to annotate qualitative
aspects such as use of the adaptive cycle, identification of its phases, whether
the paper is conceptual, modeling or an empirical study. When empirical, we
recorded the temporal and spatial scales of the case study. We also
distinguished empirical cases from empirical constructs. An empirical case is
when a study is trying to test an idea or proposition put forward by panarchy
(e.g. a cross-scale interaction, or that the back loop is faster than the forth
loop). An empirical construct is when panarchy is used as inspiration to
interpret a case study, typically with a historical component to it (e.g.
identify the phases of adaptive cycle); but it does not find support for or
against a panarchy preposition. That is, there was no mechanism in place to
avoid confirmation bias.

We also identified what aspects of panarchy were most used in the papers. For
example whether there is emphasis on cross-scale interactions, or poverty and
rigidity traps. We used text annotations to capture potential criticisms as well
as the methods used. We did a preliminary scoping and coding of papers, and
through an iterative process we kept track of potential fields under review by
our corpus, or simply absent from it. We coded papers until we reached conceptual
saturation, that is, additional papers were not proving additional categories. The code book is available at
[doi.org/10.6084/m9.figshare.13490919](http://doi.org/10.6084/m9.figshare.13490919).

# Results

Panarchy, the book published in 2002, has been cited over 8600 times according
to Google Scholar (by March 2022). The scientific paper that introduced the book to the
scientific community [@Holling:2001cu] has received, at the time of writing,
1853 citations in the Web of Science, and 2197 in Scopus. Roughly half of the
citations have come from environmental (28%) and social sciences (22%). Computer
science (2.1%) and arts and humanities (2.6%) have been the least represented.

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 5in, height = 4in]{fig2_topicnumber}
\caption{\textbf{Algorithm and number of topic selection} Gibbs sampling maximizes entropy and log-likelihood estimation, making it a suitable algorithm for our data (A). Increasing the number of topics (from 5-100) shows that $\alpha$ decreases, suggesting that despite the larger number of topics, a few of them suffice to describe most papers (B). Log-likelihood is maximized for 50 topics followed closely with 25 and 100. While 50 topics is marginally better. We choose 25 because the $\alpha$ parameter suggest preferring lower numbers, or a more parsimonious solution to avoid over fitting. Note that perplexity cannot be calculated for Gibbs sampling, hence the missing value in A and absence in B.}
\label{fig:algorithm}
\end{figure*}
```

Topic modelling was done using Gibbs sampling. In our comparison of alternative Latent Dirichlet Allocation algorithms Gibbs sampling provided the best fit
for the data (Fig \ref{fig:algorithm}). As a rule of thumb, an ideal method
should maximize entropy and the log-likelihood estimation while minimizing
perplexity [@Grun:2011tb]. Gibbs sampling maximised both entropy and
likelihood with our data when compared to other alternatives. The second best
was the variational-expectation maximisation algorithm (VEM) when $\alpha$ was
not set constant. $\alpha$ is a hyper parameter that weights the evenness of
topic distribution. A lower $\alpha$ than default values indicates that the
documents can be described by rather fewer topics, or that its distribution is
very uneven. In fact, we observed that increasing the number of topics from 5 to
100 topics did increase entropy at expense of reducing $\alpha$, meaning that
despite the larger number of topics, the main content of a document is still
captured by a few of them. The log-likelihood maximization stopped at around
25-50 topics. Thus, we restricted the rest of our analysis to 25 topics.

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 5in, height = 5in]{fig3_topics25}
\caption{\textbf{Panarchy topics over time} The number of papers per year (A) with a maximum of 324 in 2019. The proportion of papers per year (B) and the relative proportion of topic content per year (C) do not show strong trends for the time window with most papers (gray area in A). In (B), a reference dotted horizontal line marks the level where all topics are equally dominant. Each topic is summarized in (D) by the top 10-words that best describe them according to the posterior probability of our model fit. Fig \ref{fig:figS1} expands (B) presenting time series for each topic. All time series exclude 2021 to account for the artificial decline of papers at the end of the time series as shown in A.}
\label{fig:topics25}
\end{figure*}
```
A topic is a set of words that are ranked according to a probability that they
represent an underlying content of the document (Fig \ref{fig:topics25}). For
example, the words "resilience", "adaptive", or "capacity" have a high
probability to capture the content of topic 14. Papers early in our time series
(2001-3) have high content largely dominated by topics 14 on resilience, 9 on
resource management, and 23 on cross-scale interactions. Towards the end of the
time series, topic 7 on urban systems and 25 on local communities and knowledge
have spikes up to 9% of the content of each year (2019-2020). For comparison, if
all topics were equally represented in the content, they would have 4% in the
corpus (grey line in Fig \ref{fig:topics25}B). The topics presented lack names
for reproducibility purposes. While here we present e.g. topic 7 as the urban
topic, if a researcher replicates the analysis that person will find the same
list of words and probability distributions listed under a topic with a
different number. That is because the labeling of topics is random and we urge readers not to over interpret the word patterns found.

```{=tex}
\begin{figure*}[h]
\centering
\includegraphics[width = 6in, height = 5in]{SM_topics25_timeline}
\caption{\textbf{Proportion of papers per year per topic.} Each line represent the average proportion of content for a given year over all papers published in the same year. Fig \ref{fig:figS2} complements with the proportion of topics for each paper in our sample.}
\label{fig:figS1}
\end{figure*}
```
Despite fluctuations, most topics showed a relatively constant level of interest
over time (Fig \ref{fig:topics25}, \ref{fig:figS1}). We did not observe strong
trends, but some topics have gained a small amount of attention. For example,
topic 11 on sustainability, and topic 15 on ecosystem services appeared
consistently across time. In contrast, research on innovation (topic 18), urban
infrastructure (topic 7), and local communities (topic 13) have gained attention
in recent years. Topic 6 was an outlier, with a selection of papers ranking high
on content related to network infrastructure and performance, possibly from
engineering disciplines. It was the only topic with a set of papers that was
clearly distinct from the rest of the collection, and showed a decreasing trend
over time.

The human-coded document analysis revealed that the most common feature of
panarchy in the literature was the adaptive cycle (81%, 34 out of N = `r nrow(qual)`)
followed by cross-scale interactions (Fig \ref{fig:qual}). Poverty and rigidity
traps were less studied features in our sample (14% and 19% respectively), even
when accounting for slightly different terminology such as "lock-in". About half (52.4% or 22/42) of
the papers analyzed were conceptual papers, many without defined methods or
a clear research question. Roughly half (20/42) of the papers used panarchy as a metaphor. About 40% of papers in our sample (17/42) went a
step further and used panarchy as an empirical construct, for example by
attempting to identify the phases of the adaptive cycle (26/42). Off the 22
empirical cases, 6 were at the time scale of centuries, 7 on decades, 6 on
years, and 1 in weeks. Spatially, 4 were at a city scale, 17 regional, and 1
national. Empirical papers predominantly used qualitative methods (17)
and were primarily retrospective historical reviews.

Most of the papers reviewed (81%, 34 of N = 42) identified adaptive cycles
either through scientific methods, or as analogy or framework to make sense of a narrative. Identification of adaptive
cycles has found applications in a wide range of disciplines and research
problems, from delimitation of periods in archaeology and anthropology
[@Redman:2003bd], systemic interactions between forest-related sectors in Sweden [@moen2010interlocking], financial crises in Europe [@Castell:2020dg], to
traffic jams in China [@Zeng:2020hu]. The adaptive cycle has been used as a model to structure rich empirical information about participation in adaptive management and, based on that analysis, to develop recommendations for more flexible stakeholder engagement [@stringer2006unpacking]. The adaptive cycle has also been used to provide a more mechanistic description of the history of a variety of specific social-ecological systems, including the Galapagos Islands [@gonzalez2008rethinking], the Western Australian Agricultural Region [@allison2004resilience], the Erhai Lake watershed in southwest China [@dearing2008landscape], and the cities of Cambridge and Swansea in the UK [@simmie2010economic]. The adaptive cycle has also been used to describe more general social and ecological processes, such as lake eutrophication [@carpenter2001metaphor] and the development of regional industrial clusters [@martin2011conceptualizing].

Recent empirical tests of the
adaptive cycle used innovative types of data and methods. For example,
information transfer methods based on entropy have been suggested to approximate
relevant components of a system and the empirical proxies of the axes within
which the adaptive cycle unfolds: potential, connectedness, and resilience
[@Castell:2020dg]. Percolation methods combined with big data have been proven
useful for testing hypotheses of regime shifts in urban systems, an idea originally
proposed in panarchy [@Gunderson:2002vk]; and deriving the temporal and
spatial scales at which the adaptive cycle emerges [@Zeng:2020hu].

```{=tex}
\begin{figure*}[ht]
\centering
\includegraphics[width = 5in, height = 2.5in]{fig4_qual_results}
\caption{\textbf{Qualitative results} Document analysis was used to disentangle the different panarchy features addressed by a sub-sample of papers (N = 42). Over half of the papers are conceptual work, and most empirical papers fall into qualitative methods.}
\label{fig:qual}
\end{figure*}
```
# Discussion

We examined how the conceptual framework of panarchy has been developed and tested over the past twenty years. We did this using a topic modelling to analyze a large number of papers and a focused review of a selection of panarchy papers.  We find that panarchy has been used an interpretative tool across a wide variety of topics, but has there has been little effort to substantially operationalize or test panarchy.

Sustainability science is defined by the problems it addresses rather than the
disciplines it employs [@Anonymous:2016fv]. This goal of this paper was to explore how the ideas proposed in panarchy have evolved, to what type of problems they have been applied and found useful, and finally, what key research frontiers remain. Our topic modeling results helped
in identifying some of the problems that the community is addressing and how
they have changed over time. Some of these problems include climate change
(topic 24), urban development and disaster response (7), conservation and
landscape management trade-offs (5), natural resource management (9), food
production (3), policy tools and plans (16), community based management (13) or
knowledge and local social practices (25). We also observed topics closely
related to methods or approaches to study social-ecological systems such as
modeling (19,21), networks (6), or ecosystem services (15). Topic modeling can
be used to further investigate particular academic communities, identify key
landmark papers in certain topics, or identify the network of key collaborators
and experts.

This paper aimed to study the use of panarchy related concepts and their
evolution since the publication of the book [@Gunderson:2002vk]. Systematic literature reviews are
often restricted to a limited subset of the literature -- based on what is
readable on the time frame of a project -- of dozens to maybe hundreds of
papers. Qualitative analysis offers rich insights into the papers reviewed but
is limited by sample size in its ability to draw general conclusions. In our
analysis, the qualitative insights could be biased towards a non-random
selection of papers that aligned with our research interests or highly cited review
papers. Topic modeling enabled us to complement the analysis and include all
papers indexed by major scientific databases. It has the advantage of
reproducibility and reduces sample bias, but offers limited insights into the
dimensions of panarchy used, methods, or criticisms. By combining methods, as we did in this paper, we believe literature reviews can become more powerful.

Our choice of methods does however have some limitations. We could not directly
compare the two methods since they do not share a common categorical axis that we could use to translate between methods. Topic modelling contributes understanding of what topics --sets of words-- panarchy has been associated with; while our document analysis contribute in-depth understanding on how people have interpreted and use these concepts. The topics gave us a broad but
shallow overview, while the document analysis provides deep insights on
particular work but limited generalizations. Our method also ignored rich
areas of social science research (e.g. Anthropology) often published in books
and book chapters, that are not archived by scientific search engines.
Finally, an iterative inductive approach to the qualitative coding could have
given us different insights on the literature outside the pre-defined categories
we initially shared with the Resilience Alliance.

Future studies of panarchy, or other concepts, can benefit from our experience. Panarchy is a collection of concepts, some of them more used and developed than others such as the adaptive cycle. Future reviews would benefit from including gray literature such as theses,
books, non-governmental organizations reports, government agencies reports, or
non-English literature. This would let us see the usage of panarchy concepts in
more applied settings such as management and decision making. The use of full text instead of abstracts can also offer
additional insights on the automated analyses, for example in attributing
impacts of ecosystem services from regime shifts in social-ecological systems
[@Rocha:2015tk].  While databases are improving, full-text analysis are only possible for open-access journals, and text analysis is difficult across multiple languages. Databases are increasing their coverage of books and reports, which suggest our approach may be able to access a broader set of literature in the future.

Our review of panarchy papers (N = `r nrow(qual)`) revealed a high level of conceptualization but lack of theory development or testing.  Conceptual frameworks, such as panarchy, have an important role as boundary objects that enable interdisciplinary dialogues. The usefulness of panarchy in this role is revealed by how it has been used to address a variety of issues in fields ranging from anthropology to engineering. Many of papers reviewed used the adaptive cycle to interpret past events, but did not engage in measurement or observation to test adaptive cycles in the present or potential future trajectories, nor other features of panarchy such as poverty or rigidity traps, or revolt and remember cross-scale interactions. There remains a lack of testable mechanistic theories to explain the dynamics of panarchy, and consequently there is a lack of methods to measure or observe it in real social-ecological systems.

The high level of conceptualization but lack of theory testing came across as one of the major limitations, given our sample of
papers reviewed (N=`r nrow(qual)`). What makes a theory a scientific theory is
that the plurality of scientific approaches (e.g. positivism, constructivism)
should have mechanisms in place to avoid confirmation bias, allowing the
emergence of alternative explanations, and identify the circumstances under
which a theory may fail [@Hansson:2017tr]. Many of the papers we read
confirmed but did not challenge the expectations that panarchy put forward,
though a few of the most recent works did falsify some of the ideas proposed
--e.g. the back loop is faster than the front one [@Castell:2020dg]. We
believe more work is needed to better understand the circumstances where
panarchy works, and when it falls short. In other words, to refine the
scientific framework into a scientific theory.

The criticism of over-conceptualization is not unique to the literature that
engages with panarchy ideas. Here by over-conceptualization we refer to the
problem of scientific jargon, where different communities use slightly different
wordings to refer to the same phenomenon; reducing the chances that lessons from
a discipline percolates to another. An example from panarchy is the endorsement
of ideas such as creative destruction (from economics) or the theory of
intermediate disturbances (from ecology). The latter has been debunked in the
ecological literature
[@Fox:2013br]. A recent review of sustainability science mapped the different
schools of thought that the discipline has developed over the last decades
[@Clark:2020jd] and reached a similar conclusion: too many conceptual
frameworks have been developed, but there have been too few empirical attempts
to test the frameworks against data and falsify hypotheses. The review also
emphasizes the problem of measurement and observation [@Clark:2020jd]. In the
context of panarchy, recent work has developed an information theory based
approach that enables the identification of adaptive cycles
[@Castell:2020dg]. The authors identify phases of the adaptive cycle in the
European financial crises and in grassland ecosystems, but fail to find support
for the difference in speed between the forward and backward loops originally
proposed in panarchy. Lack of support for the same ideas was also found in studies of urban traffic [@Zeng:2020hu].

The panarchy dimensions that received less attention in our qualitative analysis
included creative destruction, rigidity and poverty traps. This may be at least
partly due to the scope of our data: papers that have cited panarchy or
Holling's paper [@Gunderson:2002vk; @Holling:2001cu], and a relatively small
sample size (N=`r nrow(qual)`). Concepts such as creative destruction and
poverty traps precede panarchy and therefore have been theorized and empirically
grounded outside the panarchy stream of thinking. For example, the theory of
poverty traps dates back to the 1950's in economics, and has received both
theoretical development [@Bowles:2006hx] as well as empirical grounding
[@Banerjee:2015ja; @Banerjee:2012tq] that has enabled researchers and
governments to distinguish what kind of interventions are likely to reduce
poverty.  As researchers have done with poverty traps, we believe it would be fruitful to propose and evaluate potential mechanisms that could produce adaptive cycles, their cross-scale interactions
(remember and revolt), or how the hierarchical nested nature of complex systems
enhance or erode resilience in social-ecological systems.

# Conclusion

In this paper we explore how the conceptual framework of panarchy has been used, developed and tested over the past twenty years. Despite a growing body of literature, no topic appears to dominate papers using panarchy concepts. The adaptive cycle is the most widely used concept, and it has been most used to interpret histories that range from the dynamics of natural resource management, urban development, to archeological and anthropological studies that span millennia. Cross-scale dynamics and structure and traps have received less attention.  Operationalzing panarchy theory remains a challenge.  How to measure and forecast system behaviour in ways that is useful for panarchy concepts remains an open question.  Addressing these issues would be very useful by developing tools to answer critical questions about the dynamics of resilience across linked social-ecological systems. For example, how do current changes in the global food system alter its resilience, connectedness and potential.  How do identify whether a social-ecological system is entering a different phase of the adaptive cycle?  What type of ‘remember’ or ‘revolt’ dynamics are shaping the current food systems of Stockholm or Bogota?  While our survey of the literature identified a few promising pathways, developing ways to operationalize theories that explain the dynamics of resilience is vitally needed to confront the challenges of creating a sustainable, just, world.

# Acknowledgements

JCR acknowledges financial support from Formas research grant (942-2015-731),
the Stockholm Resilience Centre, and Stockholm University. We acknowledge the
feedback of the Resilience Alliance and encouragement in pursuing this project. GDP acknowledges the support of the Rasmussen foundation.

# Data availability

All data and code to replicate the analysis are available at:
<https://github.com/juanrocha/panarchy>. The code book and database are
available in the public online repository
([https://doi.org/10.6084/m9.figshare.13490919](doi.org/10.6084/m9.figshare.13490919)).

# References

::: {#refs}
:::

\pagebreak

# Supplementary Material {#sec:SM}

```{=tex}
\renewcommand\thefigure{A\arabic{figure}}
\renewcommand\thetable{A\arabic{table}}
\setcounter{table}{0}
\setcounter{figure}{0}
```
```{=tex}
\begin{figure*}[h]
\centering
\includegraphics[width = 6in, height = 5in]{sm2}
\caption{\textbf{Proportion of topics per paper.} Each dot is a paper with its respective proportion of topic per year. A loess fit (blue lines) show the lack of trends over time for most topics. The color of topics correspond to Fig \ref{fig:topics25}}
\label{fig:figS2}
\end{figure*}
```